Repairing cooperation through a cognitive intervention in the repeated Trust Game

### 2023-05-12 article.Rmd
revision QH comments by ismail

abstract: | 
  Social trust is an important building block of strong social bonds, and its absence is a risk factor for social     dysfunction. As such, interventions to foster and strengthen trust-based cooperation are highly desirable. Using the repeated Trust Game paradigm, we assess the effectiveness of a cognitive intervention aimed at repairing the potential breakdown of cooperation from a pre-programmed, one-off defection by the opponent. Over two games, participants are given the role of the trustee and face what they believe are two different players. In between games, they either receive a cognitive intervention  or not. In reality, participants face the same computerised agent in both instances, which is programmed to play according to an HMM fitted to real players data. We saw the emergence of cooperative behavior in these games, consistent with what is expected when humans interact. This substanstiates the use of these AI agents as credible human opponents whilst affording a high degree of experimental control. The intervention led to more cooperative behavior both pre and post defection by the opponent. HMM modelling of participants actions shows participants in the intervention group had a lower probability of transitioning to non cooperative states. Posterior latent state analysis also showed a  higher proportion of players best described by more cooperative latent states in the intervention condition compared to the control condition.

@@

At the core of social interaction is figuring out the goals, intentions, and decision-making process of the interaction partner. This inference is however fraught with uncertainty, as we cannot reasonably observe these features without prior knowledge of the person. Absent a history of interaction, one needs to decide whether to take the risk of trusting others. @rousseau_introduction_1998 define trust as a "psychological state comprising the intention to accept vulnerability based upon positive expectations of the intentions or behavior of another". As such, whilst trusting and being able to signal trustworthiness are necessary conditions for building and maintaining important social bonds,  the challenge of trust is that it is by construction a risky endeavour. For example, if we deem a person trustworthy, we might decide to take the risk of investing in the relationship hoping for a collaborative outcome. If we misplace our trust, this can come at a high cost to us. Not trusting others is also risky since opportunities may be foregone, and any substitution for trust, such as excessive suspicion, over-reliance on self and other maladaptive strategies are likely to add burden and costs to a person.

Evidence from the literature emphasises the importance of social trust in determining why some people fare better than others physically and mentally [@giordano_trust_2016 ; @meng_multilevel_2014]. It affects the health status of individuals through reinforcing social support networks, maintaining community norms and facilitating collective action. Research into the determinants of psychopathology has linked trust-based constructs to the emergence of mental health disorders. @fonagy_role_2014 identified epistemic trust as an important function of early attachment relationships. Epistemic trust can be defined as the trust in the authenticity and personal relevance of interpersonally transmitted knowledge. It enables individuals to learn from their social environment in dynamic social and cultural contexts. Importantly, it allows the individual receiving social information to let go of their natural self-protective vigilance, dubbed hyper-vigilance. @fonagy_mentalizing_2017 identified hyper-vigilance, or lack of "epistemic trust" which may partially be rooted in adverse childhood experiences, as a key risk factor for the emergence of multiple mental health disorders such as Borderline Personality Disorder and Anti-Social Personality Disorder.

A popular paradigm in the study of the emergence and maintenance of trust is the repeated trust game [@joyce_trust_1995]. In this game, one player takes the role of the "investor" and is provided with a fixed endowment at the start of each trial. They get to decide how much of their endowment to invest with the other player taking the role of the "trustee". The amount that is sent is tripled and the trustee gets to decide, in return, how much of the tripled amount to send back to the investor. If the investor sends a non-zero amount, they express trust in the other player for that round, as the amount they will receive back is uncertain. Assuming the trustee sends back more than the initial investment, they signal trustworthiness to the investor and both players make gains. This scenario would constitute a cooperative exchange and would be mutually beneficial. However, cooperation in this setting is prone to be broken through intentional or simply misinterpreted actions on behalf of either player [@bendor_when_1991]. In order to repair a damaged cooperative equilibrium, players need to infer that their behaviour has violated social norms and offer amends through generous actions at potentially a high cost to them. @king-casas_rupture_2008 links the breakdown of cooperation in trustees with Borderline Personality Disorder to the absence or reduction of activation in brain regions associated with the perception of norm violations.

Since a lack of epistemic trust is a risk factor for social dysfunction [@fonagy_mentalizing_2017], and given the importance of trust for building and maintaining strong social bonds, interventions to foster and strengthen trust-based cooperation would be highly beneficial to society. Such an intervention would allow people to more easily repair broken relationships, and continue harvesting the benefits of cooperation even in the presence of accidental or intentional social norm violations. In the context of social dilemmas, there have been many attempts at fostering cooperative outcomes through either mechanism design, environment modifications or other.

Mechanism design, pioneered by @vickrey_counterspeculation_1961, is a field that studies incentive alignment and looks for ways to promote social welfare or revenue maximization, despite self-interest of the individual actors. In the context of social dilemmas, key mechanisms were explored with good effect to foster and maintain the cooperative outcome. @axelrod_evolutionary_1986 studied the emergence and stability of behavioral norms to regulate non-cooperative actions in social settings. He identifies a behavioral norm as a dominant behavioural strategy which is often punished by others when not adhered to. A meta-norm is the propensity to incur a cost in order to enforce a norm. @axelrod_evolutionary_1986 showed that when playing a social dilemma game, individuals have a strong incentive to enforce punishment of defectors lest they are in turn punished by others. This led to a decline of defection. Thus, meta-norms can be seen as a mechanism to promote and sustain cooperation in a population. In the context of the trust game, @charness_investment_2008 explored the effect a third party monitor can have on the amounts sent and received. This third-party's payoff is unaffected by the decisions made by the investor and trustee. However, the study allowed the third party to punish overly selfish trustees or reward Investors making a loss on the interaction. They found that the actions of both players were materially more cooperative in the presence of this third party. @fiedler_effect_2017 found that the introduction of a third party that monitors the investments and returns of the players led to more cooperative behavior, even when the third party had no ability to reward or punish the players.

Rather than try to modify the mechanism of the game, other approaches focused on intervening directly on the participants. @drazkowski_gratitude_2017 investigated the effect of expressing gratitude on later behaviour in the trust game. The gratitude intervention consisted of thinking about and writing down 5 things that the participants were grateful for in their life. The exercise was repeated over three sessions. Those completing the gratitude intervention reported more positive emotions and this mediated an increase in their investment in one-shot trust games. @burnham_friend-or-foe_2000 primed participants by introducing the investor as either a "friend" or "foe". They found that this priming produces significant differences in both trust (measured by the proportion of endowment sent) and trustworthiness  (measured by the proportion returned) with participants in the "friend" treatment exhibiting over twice as much trustworthiness than those in the "foe" condition.

@@

<!-- why this particular intervention? Why use this intervention for repair of trust? One motivation which you don't mention is the motivation to use cognitive probes to understand mechanisms of psychotherapy itself - this may be worth introducing.  -->

In devising potential interventions to repair trust, we can derive inspiration from the cognitive interventions championed by successful psychological therapies that aim to improve aspects of interpersonal dysfunction in BPD patients. Whilst there is no proven pharmacological therapy for BPD, some forms of psychotherapy such as Mentalisation Based Therapy (MBT) and Dialectical Behavior Therapy (DBT), have been clinically validated as efficacious approaches to improve various dysfunctional behaviors in BPD patients, including those related to social interaction [@gunderson_borderline_2018]. However, these therapies suffer from a high variability in treatment responses. Determining which interventions are effective for particular patients has been very challenging  [@rudge_mechanisms_2020; @arch_longitudinal_2012]. One promising approach is the study of how specific psychotherapeutic treatment components affect quantitative markers of behavior such as those inferred through computational modelling techniques [@huys_computational_2016]. As such, using specific cognitive probes inspired by therapeutic interventions complemented by computational modelling of behavior in tasks, may allow us to uncover cognitive mechanisms targeted by common forms of psychotherapy, as well as evaluate whether knowledge about these different mechanisms can improve the targeting of existing psychotherapies to different individuals.

In this study, we assess the effectiveness of a cognitive intervention aimed at repairing the potential breakdown of cooperation from a pre-programmed, one-off low investment sent by the investor. The intervention focuses on explaining the potential harm from reciprocating non-cooperative actions and suggesting a non-impulsive course of action to coax the investor back into cooperation. It explicitly articulates the advantages of adopting a coaxing behavior in this setting, and as such, is directly relevant to the task at hand and is very clear about the desired behavior without being overly directive. The idea being that if participants show no response to this type of direct intervention in this setting, then other forms of intervention that may target general processes such as better emotional regulation or better mentalising are then even less likely to show a benefit. As such this intervention can be seen as a litmus test for the success of more "general" forms of cognitive interventions. Moreover, in the event of the intervention's failure, analysis of the resistance to an explicit intervention itself might be worthy of further investigation to understand the mechanism behind the lack of success.

In this experiment, participants are given the role of the trustee and are randomly assigned to either a control group or an intervention group. They play two instances of the repeated Trust Game facing what they believe are two different players. In between games, they either receive a cognitive intervention (intervention condition) or not (control condition, where participants are asked to solve anagrams). In reality, participants face the same computerised agent in both instances, which is programmed to play according to an HMM fitted to real players data. We explore whether the intervention has an effect on the behaviour of the trustee and whether the learning is transferred to a new repeated Prisoner's Dilemma game when facing a seemingly new player.

<!-- DATA WRANGLING  -->
```{r dataLoad, include=FALSE }
d_finished <- readRDS("data/d_finished.RDS")
```

<!-- Before you introduce the tasks, you need to introduce the overall design of the experiment, i.e. the before/after randomized controlled experimental setup. A figure to illustrate this is a good idea. You could then have a section called 'measures' which includes details about each of the tasks (probably including the HMM), questionnaires and also emotion grid, and a section called 'interventions', which contains both the low return and the actual intervention. -->

@@

## Participants and Design
A total of 318  participants were recruited on the Prolific Academic platform. The mean age of participants was `r round(mean(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years. Participants were paid a fixed fee of £5 plus a bonus dependent on their performance. The experiment had a 2 (Condition: Intervention or Control) by 2 (Game : Pre or Post Intervention) design, with repeated measures on the first factor. Participants were randomly assigned to one of the two levels of the second factor.

@@

```{r timeline,fig.cap = "Overview of the timeline of the experiment detailing the various phases participants go through.", fig.align='center', out.width="80%"}

@@

Figure \ref{fig:timeline} shows the timeline of the experiment. After reviewing an information sheet about the experiment, participants were informed that there would be three phases to go through, with them facing a different opponent in each phase. They would face the same player throughout the first phase, consisting of the "Investment Game" with several rounds. This was the moniker used for the repeated trust game, and a total of 15 rounds were played.

After reading instructions on how to play the trust game, players were told that they would be assigned the role of the trustee.  A comprehension quiz on the trust game followed the instructions. The first trust game then ensued:  Each round started with an investment stage where the investor decided how much to send to the trustee and that choice was revealed. Immediately afterwards, participants were presented with a two-dimensional field to indicate their emotional state on two dimensions: valence, from unpleasant to pleasant (on the horizontal axis) and arousal, from low arousal to high arousal (on the vertical axis). After selecting a point on the field, participants were asked to decide how much to send back to the investor using a slider ranging from sending back nothing to sending back the whole (tripled) investment. Following that choice, an outcome page summarised both players decisions and showed the total pay-off for each player and the round then concluded. At the end of the game, participants were asked to rate the player they were facing on 4 attributes using a scale from 1 to 10: Trustworthiness, friendliness, cooperativeness and selfishness.

After being randomly assigned to either a control or intervention condition, participants were informed they would be paired with a different player in a second phase of the experiment. This second phase was similar to the first phase in terms of the game played (RTG) and the number of rounds. Finally, players were told they would be paired with a new player in the third phase of the experiment. This third phase consisted of 7 rounds of the prisoner's dilemma game, facing the new player. Throughout all games, players were explicitly told to aim to maximise the number of points as their bonus would depend on the score they accumulated throughout the experiment. The total number of rounds was not communicated to the participants in any of the games played.

After the three phases were completed, participants were asked to complete a series of questionnaires at the end of the experiment. These included: the PAI-BOR measuring Borderline traits [@morey_personality_1991], the DERS measuring Emotion Regulation ability [@gratz_multidimensional_2004] and the RFQ8 for mentalising abilities [@fonagy_development_2016]. Finally, participants were asked a series of questions around how they played each phase. First, participants were asked whether they thought they played differently in the second compared to the first phase of the RTG. Second, participants had to select whether they thought the opponents they faced were human or computer agents. Finally, we revealed to the participants that they faced the same computer agent throughout the RTG. We explained that any change in their perception of the opponent or any change in how the opponent played is due to their own change in behavior as the agent was simply reacting to their actions. Finally, we asked participants to reveal whether this experiment has taught them anything about how they should behave in social situations.

```{r interventionSlide,fig.cap = "Screenshot of the main  slide in the intervention condition", fig.align='center', out.width="80%"du}

knitr::include_graphics("figures/intervention_text.png")

@@

After a total of 15 rounds of the trust game, players were given either an "intervention" or a "control" manipulation. The intervention consisted of presenting a hypothetical scenario in which they were playing the repeated trust game and the investor would send a low investment in a new round after having previously sent higher amounts. Participants were then asked how they would react in this situation and what sort of return (low or high) they were thinking of sending back. The players then were presented with an educational slide about the benefits of not resorting to impulsive decisions such as punishment when they feel they have been wronged. In this text, players were told that punishment can create a negative feedback loop where the other player might trust them even less. An alternative action was suggested, whereby players would respond kindly to such a transgression in the hope of gaining trust from the investor. The full text of the intervention slide is presented in Figure \ref{fig:interventionSlide}. Afterwards, participants were asked whether they would send a low or high return in the same hypothetical scenario now that they have read the information on the slide. Players were asked to justify their answer. For each question during this intervention, participants had to wait for a fixed duration of 20 seconds before being able to write their answers, and they were prevented from proceeding before that time was up.  This choice was made to allow participants to engage with the questions, think about their answers and provide meaningful feedback.

@@

## Measures

### Repeated Trust Game 
Participants played two iterations of the standard version of the repeated trust game [@joyce_trust_1995]. The game is played in dyads, with one player assigned the role of the "investor" and the other player that of the "trustee". In the variant of the game we chose, the investor is endowed with 20 units at the start of each round. They need to decide how much of that endowment they want to invest with the trustee. The investment is then multiplied by a factor of 3 and sent to the trustee. As shown in Figure \ref{fig:trustGame}, the trustee in turn needs to decide how much of the (multiplied) investment they want to send back to the investor. If they send back more than a third of what they received, then both the investor and the trustee make a gain. Otherwise, the trustee would benefit but the investor would lose on their investment. If the interaction lasts for only one round, the Nash equilibrium of this game is for the investor to send nothing, as there is no incentive for the trustee to send back any return. In the repeated version, to maximise their rewards over time, both players need to build trust so that they can share the benefits of the bigger pie (the investment multiplied by three). If during the course of the interaction, the investor has been rewarded for taking the risk of sending an investment, then they would be more likely to invest more, and the multiplied investment would give both a bigger pie to share. If however the investor felt that they were not getting a return on their investment, then they would refrain from investing and neither would gain from the interaction.

```{r trustGame,fig.cap = "Screenshot of the repeated Trust Game. The game is played in dyads, with one player assigned the role of the investor and the other player that of the trustee. The investor is endowed with 20 units at the start of each round. They need to decide how much of that endowment they want to invest with the trustee. The investment is then multiplied by a factor of 3 and sent to the trustee. The trustee in turn needs to decide how much of the multiplied investment they want to send back to the investor. Shown here is the stage at which the trustee makes a decision of how much to send back the investor", fig.align='center'}

knitr::include_graphics("figures/trust_game.png")

### Repeated Prisoner's Dilemma 
The other game the participants played is the repeated Prisoner's Dilemma (RPD). Over multiple rounds, participants could choose one of two actions: A cooperative action that would yield a high pay-off if the other person also cooperated, and the lowest possible pay-off if they did not cooperate. Or a non-cooperative option that would yield a high pay-off if the other person chooses the cooperative action and a lower pay-off if they also defect. Figure \ref{fig:payoffsRPD} shows the payoff of each combination of actions as presented to the participants.

```{r payoffsRPD,fig.cap = "Screenshot of the Repeated Prisoner's Dilemma game. Over multiple rounds, participants could choose one of two actions: A cooperative action that would yield a high pay-off if the other person also cooperated, and the lowest possible pay-off if they did not cooperate. Or a non-cooperative option that would yield a high pay-off if the other person chooses the cooperative action and a lower pay-off if they also defect. Shown here is the table explaining the payoffs of each combination of actions the participant and their opponent choose", fig.align='center'}

knitr::include_graphics("figures/screenshot_RPD_payoffs.png")

### Feedback post investment in the repeated Trust Game

```{r grid,fig.cap = "Screenshot of the two axis grid in the Coaxing condition where participants were asked to report the valence and arousal of their emotional response ", fig.align='center'}

knitr::include_graphics("figures/emotion_grid.png")

Whilst playing the RTG, participants were asked to provide some feedback, in each round of the game, after seeing the amount sent by the investor. This feedback took the form of the answer to a question using a two dimensional field. What the axes of the field represented differed by condition. In the intervention condition, participants were asked to rate their emotional state with regard to the other player's choice of investment, with the x-axis representing the valence of the emotion they felt (positive or negative) and the Y-axis representing the emotional "arousal" they experienced (low or high). In the control condition, players were asked to rate attributes of the investment that were not related to their emotion, with the X-axis rating of how fast the investor is (slow or fast) and the Y-axis the magnitude of the investment (low or high). Figure \ref{fig:grid} shows a screenshot of the field in the intervention condition where participants were asked about their own emotional response to the investment. At the beginning of each game, participants were provided with detailed explanations of the meaning of the two axes as well as the opportunity to provide a baseline emotional state through using the field prior to the start of the game.

### Design of the AI agents in the RTG and RPD

```{r}
Anti_social <-  c(0.1025436430408, 0.1221931236853, 0.1345215238995,
0.1368182252617, 0.1285592471751, 0.1116014322677, 0.0895041220882,
0.0663166721334, 0.0453950277118, 0.0287077419587, 0.0167723588011,
0.0090530028158, 0.0045143317507, 0.0020796744990, 0.0008851094702,
0.0003480141146, 0.0001264134789, 0.0000424213859, 0.0000131513231,
0.0000037665630, 0.0000009965755)

@@

```{r invHMMPlot, fig.cap="Distribution of investments by the artificial investor agent conditional on its latent state as estimated by a three state hidden Markov model with a discretised truncated Gaussian as a response function",fig.align="center", fig.width=6, fig.height = 4}

ggplot(response_probs,                            
       aes(x = investment,
           y = probability,
           fill = Investor_state)) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  labs(fill='Latent investor state') + 
  theme_bw() + 
  theme(legend.position = "bottom")

@@

```{r invHmmTrans, fig.cap="Transition function for investor conditional on current state",fig.align="center", fig.width=6, fig.height = 4}

plot_HMM_transitions(3, pars_inv)

@@

The role of the investor was in reality played by an adaptive artificial agent whose strategy was modelled on the behaviour of human participants taking the role of the investor in the RTG. Using data from human dyadic interaction in the 10 round trust game, we estimate a Hidden Markov Model (HMM) to characterise the investor's behavior as emanating from a small number of latent states [@rabiner_hmm_1989]. In this instance, for both iterations of the repeated trust game, the HMM used was identical: it had three states that can be described as "low-trust", "medium-trust" and "high-trust". Each latent state was associated with a distribution over all possible investor actions (from 0 to 20 investment) that reflected the amount of trust, as presented in Figure \ref{fig:invHMMPlot}.

<!-- The advantages of this approach is that it does not require any a priori assumptions about the model features. The number of states, the policy conditional on the state, and the transition function between states can all determined in a purely data driven way. These HMMs can in turn be used to simulate a human-like agent playing the trust game. This agent may transition to a new state depending on the other player's actions and adopt a policy reflecting its state, thus simulating changes in emotional dispositions of human players during a repeated game. When the investor gains from the interaction, they become more likely to transition to a state where their policy is to be more trusting and invest more. However, faced with losses, the investor is more likely to transition to a more cautious policy where its investment belies lower trust in the trustee. The policies and the transitions between states are sufficient to build an agent that reflects this type of adaptive behavior and reacts to the trustee's action choices in a way that mimics a human player.  -->

The HMM assumes that the probability of each investment $I_t=0,…,20$, at each trial t, conditional on the current state of the investor $S_t$, is dependent on an underlying normal distribution with mean $\mu_s$ and standard deviation $\sigma_s$. The probability of each discrete investment was determined by the cumulative normal distribution $\Phi$. For instance, the probability of an investment $I_t=2$ is defined as:

>>>>

abstract: | 
  Social trust is an important building block of strong social bonds, and its absence is a risk factor for social dysfunction. As such, interventions to foster and strengthen trust-based cooperation are highly desirable. Using the repeated Trust Game paradigm, we assess the effectiveness of a cognitive intervention aimed at repairing the potential breakdown of cooperation from a pre-programmed, one-off defection by the opponent. Over two games, participants are given the role of the trustee and face what they believe are two different players. In between games, they either receive a brief cognitive intervention or not. The intervention led to more cooperative behavior both pre and post defection by the opponent. HMM modelling of participants actions shows participants in the intervention group had a lower probability of transitioning to non cooperative states. Posterior latent state analysis also showed a  higher proportion of players best described by more cooperative latent states in the intervention condition compared to the control condition.

@@

At the core of social interaction is figuring out the goals, intentions, and decision-making process of the interaction partner. This inference is however fraught with uncertainty, as we cannot reasonably observe these features without prior knowledge of the person. Absent a history of interaction, one needs to decide whether to take the risk of trusting others. The challenge of trust is that it is by construction a risky endeavour. For example, if we deem a person trustworthy, we might decide to take the risk of investing in the relationship hoping for a collaborative outcome. If we misplace our trust, this can come at a high cost to us. Not trusting others is also risky since opportunities for cooperation may be foregone.

<!-- and any substitution for trust, such as excessive suspicion, over-reliance on self and other maladaptive strategies are likely to add burden and costs to a person. -->

Evidence from the literature emphasises the importance of social trust in determining why some people fare better than others physically and mentally [@giordano_trust_2016 ; @meng_multilevel_2014]. It affects the health status of individuals through reinforcing social support networks, maintaining community norms and facilitating collective action. Research into the determinants of psychopathology has linked trust-based constructs to the emergence of mental health disorders. @fonagy_role_2014 identified epistemic trust, or the belief in the authenticity and personal relevance of interpersonally transmitted knowledge, as an important function of early attachment relationships. It allows the individual receiving social information to let go of their natural self-protective vigilance, which can become a pathological hypervigilance frequently observed after traumatic experiences and a key factor for the emergence of multiple mental health disorders [@fonagy_mentalizing_2017].

<!-- A well-established paradigm in the study of the emergence and maintenance of trust is the repeated trust game [@joyce_trust_1995]. In this game, one player takes the role of the "investor" and is provided with a fixed endowment at the start of each trial. They get to decide how much of their endowment to invest with the other player taking the role of the "trustee". The amount that is sent is tripled and the trustee gets to decide, in return, how much of the tripled amount to send back to the investor. If the investor sends a non-zero amount, they express trust in the other player for that round, as the amount they will receive back is uncertain. Assuming the trustee sends back more than the initial investment, they signal trustworthiness to the investor and both players make gains. This scenario would constitute a cooperative exchange and would be mutually beneficial. However, cooperation in this setting is prone to be broken through intentional or simply misinterpreted actions that can convey lack of trust or trustworthiness such as low investments and returns [@bendor_when_1991]. In order to repair a damaged cooperative equilibrium, players need to infer that their behaviour has violated social norms and offer amends through generous actions at potentially a high cost to them. @king-casas_rupture_2008 links the breakdown of cooperation in trustees with Borderline Personality Disorder to the absence or reduction of activation in brain regions associated with the perception of norm violations.  -->

Since a lack of epistemic trust is a risk factor for social dysfunction [@fonagy_mentalizing_2017], and given the importance of trust for building and maintaining strong social bonds, interventions to foster and strengthen trust-based cooperation would be highly beneficial to society. Such interventions would allow people to more easily repair broken relationships, and continue harvesting the benefits of cooperation even in the presence of accidental or intentional social norm violations.

<!-- Mechanism design, pioneered by @vickrey_counterspeculation_1961, is a field that studies incentive alignment and looks for ways to promote social welfare or revenue maximization, despite self-interest of the individual actors. In the context of social dilemmas, key mechanisms were explored with good effect to foster and maintain the cooperative outcome. @axelrod_evolutionary_1986 studied the emergence and stability of behavioral norms to regulate non-cooperative actions in social settings. He identifies a behavioral norm as a dominant behavioural strategy which is often punished by others when not adhered to. A meta-norm is the propensity to incur a cost in order to enforce a norm. @axelrod_evolutionary_1986 showed that when playing a social dilemma game, individuals have a strong incentive to enforce punishment of defectors lest they are in turn punished by others. This led to a decline of defection. Thus, meta-norms can be seen as a mechanism to promote and sustain cooperation in a population. In the context of the trust game, @charness_investment_2008 explored the effect a third party monitor can have on the amounts sent and received. This third-party's payoff is unaffected by the decisions made by the investor and trustee. However, the study allowed the third party to punish overly selfish trustees or reward Investors making a loss on the interaction. They found that the actions of both players were materially more cooperative in the presence of this third party. @fiedler_effect_2017 found that the introduction of a third party that monitors the investments and returns of the players led to more cooperative behavior, even when the third party had no ability to reward or punish the players.  -->

A well-established paradigm in the study of trust is the repeated trust game [@joyce_trust_1995]. To encourage the emergence and maintenance of trust in this setting, some studies focused on modifying the game mechanism [@charness_investment_2008; @fiedler_effect_2017]. Others chose to intervene directly on the participants. For example, @drazkowski_gratitude_2017 found that trust was increased when participants were asked to think about and write down five things that they were grateful for. @burnham_friend-or-foe_2000 found that trust was also increased when participants were primed with the concepts of friend and foe.

@@

In devising potential interventions to repair trust, we can derive inspiration from the cognitive interventions championed by successful psychological therapies that aim to improve aspects of interpersonal dysfunction in BPD patients. Whilst there is no proven pharmacological therapy for BPD, some forms of psychotherapy such as Mentalisation Based Therapy [@allen_handbook_2006-1] and Dialectical Behavior Therapy [@linehan_cognitive-behavioral_1993], have been clinically validated as efficacious approaches to improve various dysfunctional behaviors in BPD patients, including those related to social interaction [@gunderson_borderline_2018]. However, these therapies suffer from a high variability in treatment responses. Determining which interventions are effective for particular patients has been very challenging [@rudge_mechanisms_2020; @arch_longitudinal_2012]. One promising approach is the study of how specific psychotherapeutic treatment components affect quantitative markers of behavior such as those inferred through computational modelling techniques [@huys_computational_2016; @reiter_neuro-cognitive_2021]. As such, using specific cognitive probes inspired by therapeutic interventions complemented by computational modelling of behavior in tasks, may allow us to uncover cognitive mechanisms targeted by common forms of psychotherapy. This may in turn form the basis for targeting of existing psychotherapies to different individuals.

In this study, we assess the effectiveness of a cognitive intervention aimed at repairing the potential breakdown of cooperation from a pre-programmed, one-off low investment sent by an computer investor. The intervention focuses on explaining the potential harm from reciprocating non-cooperative actions and suggesting a non-impulsive course of action to coax the investor back into cooperation. Participants are given the role of the trustee and are randomly assigned to either a control group or an intervention group. They play two instances of the repeated Trust Game facing what they believe are two different players. In between games, they either receive a cognitive intervention (intervention condition) or not (control condition, where participants are asked to solve anagrams). In reality, participants face the same computerised agent in both instances, which is programmed to play according to an HMM fitted to real players data. We explore whether the intervention has an effect on the behaviour of the trustee and whether the learning is transferred to a new repeated Prisoner's Dilemma game when facing a seemingly new player.

```{r dataLoad, include=FALSE }
d_finished <- readRDS("data/d_finished.RDS")
```

@@

## Participants and Design

A total of 318 participants were recruited on the Prolific Academic platform (prolific.co). The mean age of participants was `r round(mean(as.numeric(d_finished$exitSurvey.age), na.rm=TRUE),1)` years. Participants were paid a fixed fee of £5 plus a bonus dependent on their performance. The experiment had a 2 (Condition: Intervention or Control) by 2 (Game : Pre or Post Intervention) design, with repeated measures on the first factor. Participants were randomly assigned to one of the two levels of the second factor.

@@

```{r timeline, fig.cap = "A) Experiment overview. After playing 15 rounds of a repeated trust game (RTG) as trustee, participants were randomized to either receive the control or active intervention. They then played the second set of 15 rounds of RTG (again as trustee) to examine intervention effects, and 7 rounds of a repeated prisoner’s dilemma to examine generalization of intervention effects. Finally, participants answered questionnaires and were debriefed.", fig.align='center', out.width="80%"}

@@

<!-- Figure \ref{fig:timeline} shows the timeline of the experiment. After reviewing an information sheet about the experiment, participants were informed that there would be three phases to go through, with them facing a different opponent in each phase. They would face the same player throughout the first phase, consisting of the "Investment Game" with several rounds. This was the moniker used for the repeated trust game, and a total of 15 rounds were played. Participants were then randomly assigned to either a control or intervention condition. -->

<!-- <!-- After reading instructions on how to play the trust game, players were told that they would be assigned the role of the trustee.  A comprehension quiz on the trust game followed the instructions. The first trust game then ensued:  Each round started with an investment stage where the investor decided how much to send to the trustee and that choice was revealed. Immediately afterwards, participants were presented with a two-dimensional field to indicate their emotional state on two dimensions: valence, from unpleasant to pleasant (on the horizontal axis) and arousal, from low arousal to high arousal (on the vertical axis). After selecting a point on the field, participants were asked to decide how much to send back to the investor using a slider ranging from sending back nothing to sending back the whole (tripled) investment. Following that choice, an outcome page summarised both players decisions and showed the total pay-off for each player and the round then concluded. At the end of the game, participants were asked to rate the player they were facing on 4 attributes using a scale from 1 to 10: Trustworthiness, friendliness, cooperativeness and selfishness. -->

<!-- In the second phase of the experiment, participants were informed they would be paired with a different player. This phase was similar to the first phase in terms of the game played (RTG) and the number of rounds. In all rounds of the RTG, participants gave feedback after seeing the investment received. This feedback was given though a two dimensional grid that was either related to their emotional state (degree of valence of arousal) in the Intervention group, or unrelated to the emotional state in the control group (speed of investor's response and magnitude of investment). At the end of each RTG, participants were asked to rate how cooperative, selfish, trustworthy and friendly they thought their interaction partner was. Details of the two dimensional grid and player ratings are given in the supplement. -->

<!-- After both trust games,  players were told they would be paired with a new player in the third phase of the experiment. This third phase consisted of 7 rounds of the prisoner's dilemma game, facing the new player. Throughout all games, players were explicitly told to aim to maximise the number of points as their bonus would depend on the score they accumulated throughout the experiment. The total number of rounds was not communicated to the participants in any of the games played. -->

<!-- After the three phases were completed, participants were asked to complete a series of questionnaires at the end of the experiment.  -->

<!-- <!-- These included: the PAI-BOR measuring Borderline traits [@morey_personality_1991], the DERS measuring Emotion Regulation ability [@gratz_multidimensional_2004] and the RFQ8 for mentalising abilities [@fonagy_development_2016].  -->

<!-- Finally, participants were asked a series of questions around how they played each phase. First, they were asked whether they thought they played differently in the second compared to the first phase of the RTG. Second, participants had to select whether they thought the opponents they faced were human or computer agents. Finally, we revealed to the participants that they faced the same computer agent throughout the RTG. We explained that any change in their perception of the opponent or any change in how the opponent played is due to their own change in behavior as the agent was simply reacting to their actions. Finally, we asked participants to reveal whether this experiment has taught them anything about how they should behave in social situations. -->

The experiment as shown in Figure \ref{fig:timeline} began with participants briefed there will be three phases, and that they would face a new opponent in each phase. Phase one was a 15 round Trust Game (RTG) in which participants had the role of the trustee, facing the same investor opponent. Participants were then assigned to either the control or intervention conditions. Phase two was similar to Phase one, with participants being told they faced a new player. In both phases, participants were asked for feedback in each round after seeing the investment. The nature of the feedback solicited varied by condition, either emotion-related (intervention) or unrelated (control). Participants rated their interaction partner on various attributes after each phase. More details on these attributes, nature of feedback and exact instructions are found in the supplement. Phase three involved 7 rounds of the repeated prisoner's dilemma game (RPD) with a third player. Participants were explicitly told to aim to maximise the number of points in all phases. The total number of rounds was not communicated to the participants in any of the games played.

Post-experiment, participants completed questionnaires related to mentalising abilities, emotion regulation and BPD traits (see details in supplement). They were then quizzed on their game-play, including whether they detected that all opponents were computer agents. The reveal explained that they were facing identical AI agents and that any perceived opponent changes were due to participants' own behavior.

@@

The intervention was built on interventions from DBT skills training, asking patients to reflect on the consequences of actions taking in emotional states [@linehan_dbt_2015]. Specifically, participants were presented with a low investment and asked to indicate their response. They were then invited to consider what their ultimate aim in the game was and whether this response was most likely to achieve their aim. The intervention is detailed in the supplementary information.

@@

## Design of the AI agents in the RTG and RPD

```{r, include=FALSE}
Anti_social <-  c(0.1025436430408, 0.1221931236853, 0.1345215238995,
0.1368182252617, 0.1285592471751, 0.1116014322677, 0.0895041220882,
0.0663166721334, 0.0453950277118, 0.0287077419587, 0.0167723588011,
0.0090530028158, 0.0045143317507, 0.0020796744990, 0.0008851094702,
0.0003480141146, 0.0001264134789, 0.0000424213859, 0.0000131513231,
0.0000037665630, 0.0000009965755)

@@

plotinvHMM <- ggplot(response_probs,                            
       aes(x = investment,
           y = probability,
           fill = Investor_state)) +
  geom_bar(stat = "identity",
           position = "dodge") + 
  labs(fill='Latent investor state') + 
  theme_bw() + 
  theme(legend.position = "bottom")

@@

plotInvTran <- plot_HMM_transitions(3, pars_inv)

@@

```{r HMMPanels, fig.cap="A: Distribution of investments by the artificial investor agent conditional on its latent state as estimated by a three state hidden Markov model fitted to human dyadic play dataset. B: Transition function for the HMM investor conditional on its current state and the net return from the previous round of play.", fig.align="center"}

ggpubr::ggarrange(plotinvHMM , plotInvTran, nrow=2, ncol=1, labels = c("A", "B"), common.legend = TRUE, legend = "bottom")

```
The role of the investor was in reality played by an adaptive artificial agent whose strategy was modelled on the behaviour of human participants taking the role of the investor in the RTG. Using data from human dyadic interaction in the 10 round trust game, we estimate a Hidden Markov Model (HMM) to characterise the investor's behavior as emanating from a small number of latent states [@rabiner_hmm_1989]. In this instance, for both iterations of the repeated trust game, the HMM used was identical: it had three states that can be described as "low-trust", "medium-trust" and "high-trust". Each latent state was associated with a distribution over all possible investor actions (from 0 to 20 investment) that reflected the amount of trust, as presented in Figure \ref{fig:HMMPanels}.A. Transition probabilities between these latent states were modelled as a function of the net return in the previous round conditional on the investor's state as shown in Figure \ref{fig:HMMPanels}.B.
